Commutative character sheaves and geometric types for supercuspidal representations

### 2017-06-28 Geometric types.tex
Working on PGL2/SL2 examples
Push before running off ...

%%%%%%%%%%%% MISCELLANEOUS OPERATORS %%%%%%%%%%%%
\newcommand{\sheafHom}{{\mathscr{H}\hskip-4pt{\it o}\hskip-2pt{\it m}}}
\newcommand{\abs}[1]{{\vert #1 \vert}}
\newcommand{\ceq}{{\, :=\, }}
\newcommand{\tq}{{\ \vert\ }}
\newcommand{\iso}{{\ \cong\ }}
\newcommand{\trFrob}[1]{t_{#1}}
\DeclareMathOperator{\Tr}{Tr}
\newcommand{\TrFrob}[1]{\Tr_{#1}}
%% Limits
\newcommand{\invlim}[1]{\lim\limits_{\overleftarrow{#1}}}
\newcommand{\dirlim}[1]{\lim\limits_{\overrightarrow{#1}}}
\newcommand{\limit}[1]{\mathop{\textsc{lim}}\limits_{#1}}
\newcommand{\colimit}[1]{\mathop{\textsc{colim}}\limits_{#1}}
%% Fonts for quasicharacter sheaves
\newcommand{\cs}[1]{{\mathcal{#1}}}
\newcommand{\gcs}[1]{{\mathcal{\bar #1}}}
\newcommand{\dualgcs}[1]{\gcs{#1}^\dagger}
\newcommand{\dualcs}[1]{\cs{#1}^\dagger}
%% Categories
\newcommand{\CS}{{\mathcal{C\hskip-0.8pt S}}}
\newcommand{\CCS}{{\mathcal{C\hskip-.8pt C\hskip-0.8pt S}}}
\newcommand{\bCS}{{\CS_0}}
\newcommand{\CSab}{\CS_{\operatorname{ab}}}
\newcommand{\catname}[1]{\normalfont{\textsf{#1}}}
\newcommand{\Sch}[1]{{\catname{Sch}_{/#1}}}
\newcommand{\QCS}{{\mathcal{QC\hskip-0.8pt S}}}
\newcommand{\CSiso}[1]{\CS(#1)_{/\text{iso}}}
\newcommand{\bCSiso}[1]{\bCS(#1)_{/\text{iso}}}
\newcommand{\QCSiso}[1]{\QCS(#1)_{/\text{iso}}}
\newcommand{\CCSiso}[1]{\CCS(#1)_{/\text{iso}}}
%% Labeled items
\makeatletter
\newcommand{\labitem}[2]{
\def\@itemlabel{\textbf{#1}}
\item
\def\@currentlabel{#1}\label{#2}}
\makeatother
%% Shorthand for bars
\renewcommand{\bf}{\bar{f}}
\newcommand{\bg}{{\bar{g}}}
\newcommand{\bm}{\bar{m}}
\newcommand{\bG}{\bar{G}}
\newcommand{\bH}{\bar{H}}
\newcommand{\brho}{{\bar\rho}}
\newcommand{\bx}{{\bar{x}}}
%% Spacing control
\newcommand{\tight}[3]{\hspace{-#1pt}{#2}\hspace{-#3pt}}
\newcommand{\GxG}{\text{$G \tight{1}{\times}{1} G$}}
\newcommand{\bGxG}{\text{$\bar{G} \tight{1}{\times}{1} \bar{G}$}}
\newcommand{\bfxf}{\text{$\bar{f} \tight{1}{\times}{1} \bar{f}$}}
\newcommand{\GxxG}{\text{$G \tight{1}{\times}{1} G$}}
\newcommand{\LxL}{\text{$\gcs{L} \tight{0}{\boxtimes}{0} \gcs{L}$}}

@@

\begin{definition}\label{def:CSab}
Let $\CSab(G)$ denote the category of triples $(\cs{L},\mu,\beta)$ where $(\cs{L},\mu) \in \CS(G)$ and
$\beta : \cs{L}\vert_{G\der} \to (\EE)_{G\der}$ is an isomorphism in $\CS(G\der)$.
A morphism $(\cs{L},\mu,\beta)\to (\cs{L}',\mu',\beta')$ is a morphism $\alpha : (\cs{L},\mu)\to (\cs{L}',\mu')$
in $\CS(G)$ such that $\beta = \beta' \circ \alpha\vert_{G\der}$.  
\end{definition}

The reason for tracking $\beta$ is that it determines an isomorphism $\gamma : m^*\cs{L} \to \theta^*m^*\cs{L}$ which will replace the $\xi$ of Definition \ref{def:CCScom}, as follows.
Let $i : G \to G$ be inversion and $c : G\times G\to G\der$ be the commutator map, defined by $c(x,y)= xyx^{-1}y^{-1}$.
Both are morphisms of $\Fq$-schemes.
Set $m' = i \circ m \circ \theta$ and let $i\der : G\der\to G$ be inclusion; then $i\der\circ c = m \circ (m \times m')$.
Then $\beta : \cs{L}\vert_{G\der} \to (\EE)_{G\der}$ determines the isomorphism $\gamma' : m^*\cs{L} \otimes \theta^* m^* i^*\cs{L} \to (\EE)_{G\times G}$ by the diagram of isomorphisms below.
\begin{equation}
\begin{tikzcd}
\arrow[equal]{d} c^* (\cs{L}\vert_{G\der}) \arrow{r}{c^*(\beta)} 
	&  c^*((\EE)_{G\der}) \arrow[equal]{d} \\
\arrow{d}[swap]{i\der\circ c = m \circ (m \times m')} c^* i\der^* \cs{L} 
	&   (\EE)_{G\times G} \\
(m \times m')^* m^* \cs{L} \arrow{d}[swap]{(m \times m')^*(\mu)} 
	&  m^*\cs{L} \otimes \theta^* m^* i^* \cs{L}  \arrow[dashed]{u}[swap]{\gamma'} \\
(m \times m')^* (\cs{L} \boxtimes \cs{L}) \arrow[equal]{r} 
	& m^*\cs{L} \otimes (m')^*\cs{L} \arrow{u}[swap]{m' = i\circ m\circ \theta} 
\end{tikzcd}
\end{equation}
In the diagram above, the arrows labeled with equations come from canonical isomorphisms of functors on Weil sheaves derived from the equations; so, for example, the middle left isomorphism comes from $(m\times m')^* m^* \iso c^* i\der^*$ since $i\der\circ c = m \circ (m \times m')$.
Using the monoidal structure of the category of Weil local systems on $G\times G$, the isomorphism $\gamma' : m^*\cs{L} \otimes \theta^* m^* i^*\cs{L} \to (\EE)_{G\times G}$ defines an isomorphism
\[
m^*\cs{L} \to (\theta^* m^* i^*\cs{L})^\vee.
\]
Applying the canonical isomorphisms $(\theta^* m^* i^*\cs{L})^\vee \iso \theta^* m^* i^* (\cs{L}^\vee)$ and $i^*(\cs{L}^\vee) \cong \cs{L}$, this map provides the promised isomorphism
\[
\begin{tikzcd}
\gamma : m^*\cs{L} \arrow{r} & \theta^* m^* \cs{L}.
\end{tikzcd}
\]

\begin{definition}\label{def:CCS}
The category $\CCS(G)$ of commutative character sheaves on $G$ is the full subcategory of $\CSab(G)$ consisting of triples $(\cs{L},\mu,\beta)$ such that the following diagram of Weil sheaves on $G \times G$ commutes:
  \[
  \begin{tikzcd}[row sep=30]
   m^*\cs{L} \arrow{d}[swap]{\gamma} \arrow{r}{\mu} & \cs{L}\boxtimes\cs{L} \arrow{d}{\vartheta}\\
   \theta^*(m^*\cs{L}) \arrow{r}{{\theta}^*\mu} &  \theta^*(\cs{L}\boxtimes\cs{L}).
  \end{tikzcd}
  \]
Here $\gamma : m^*\cs{L} \to \theta^* m^* \cs{L}$ is the isomorphism built from $\beta : \cs{L}\vert_{G\der} \to (\EE)_{G\der}$ as above. 
\end{definition}

@@

In this section we give an equivalence of categories between $\CS(G\ab)$ and $\CSab(G)$ and use it to describe
the pullback functor $q^* : \CS(G\ab) \to \CS(G)$ in terms of the forgetful functor $\CSab(G) \to \CS(G)$, where $q : G\to G\ab$ is the abelianization quotient with kernel $G\der$.
But first, in order to study commutative character sheaves, we need some auxiliary categories.

@@

Let $\Loc(G)$ be the category of Weil local systems on $G$.
Let $\Loc\der(G)$ be the category of $G\der$-equivariant Weil local systems on $G$, whose definition we now recall.
Let $n : G\der\times G\to G$ be the restriction of $m : G\times G\to G$ to $G\der\times G$,
let $p : G\der\times G \to G$ be projection to the second component, and let $s: G \to G\der\times G$ be given by $s(g) = (1,g)$.
Then the quotient $q : G \to G\ab$ is a regular epimorphism of smooth group schemes with kernel pair $(n,p)$.
\[
\begin{tikzcd}
G\der\times G
 \arrow[shift left=1]{r}{n}
  \arrow[shift right=1,swap]{r}{p}
&
G 
\arrow{r}{q}
& 
G\ab
\end{tikzcd}
\]
Consider the morphisms
\[
\begin{tikzcd}
G\der \times G\der \times G 
\arrow[shift left=2]{r}{b_1, b_2, b_3} 
\arrow{r}{}
\arrow[shift right=2]{r}{} 
& G\der \times G 
\arrow[shift left=1]{r}{n}
\arrow[shift right=1]{r}[swap]{p}
 & G 
\end{tikzcd}
\]
defined by 
\begin{align*}
b_1(h_1,h_2,g) &= (h_1h_2,g) \\
b_2(h_1,h_2,g) &= (h_1,h_2g) \\
b_3(h_1,h_2,g) &= (h_2,g).
\end{align*}
Note that
\begin{align*}
n\circ b_1 &= n\circ b_2 \\
n\circ b_3 &= p\circ b_2 \numberthis \label{eqn:bap}\\
p\circ b_1 &= p\circ b_3.
\end{align*}
A $G\der$-equivariant Weil local system on $G$ is a Weil local system $\cs{L}$ on $G$ together with an isomorphism  
\[
\nu : n^*\cs{L} \to p^*\cs{L}
\] 
of Weil local systems on $G\der\times G$ such that 
\begin{equation}\label{E1}
s^*(\nu) = \id_{\cs{L}}
\end{equation}
and such that the following diagram of isomorphisms of local systems on $G\der\times G\der\times G$ commutes. 
\begin{equation}\label{E2}
\begin{tikzcd}
\ &  \arrow{dl}{n\circ b_1 = n\circ b_2}  b_2^*  n^*\, \mathcal{L} \arrow{rr}{b_2^*(\nu)} && b_2^* p^*\, \mathcal{L} \arrow{dr}[swap]{p\circ b_2 = n\circ b_3} & \\
b_1^* n^*\, \mathcal{L} \arrow{dr}{b_1^*(\nu)} &&&&   \arrow{dl}[swap]{b_3^*(\nu)} b_3^* n^*\, \mathcal{L} \\
& b_1^* p^*\, \mathcal{L}  && \arrow{ll}[swap]{p\circ b_3 = p\circ b_1} b_3^* p^*\, \mathcal{L} & 
\end{tikzcd}
\end{equation}
Morphisms of $H$-equivariant Weil local systems $(\cs{L},\nu)\to (\cs{L}',\nu')$ are morphisms of Weil local systems $\alpha: \cs{L}\to \cs{L}'$ for which the diagram
\begin{equation}\label{E3}
\begin{tikzcd}
\arrow{d}[swap]{\nu} n^*\cs{L} \arrow{r}{n^*(\alpha)} & n^*\cs{L}' \arrow{d}{\nu'} \\
p^*\cs{L} \arrow{r}{p^*(\alpha)} & p^*\cs{L}'
\end{tikzcd}
\end{equation}
commutes.
This defines $\Loc\der(G)$, the category of $G\der$-equivariant Weil local systems on $G$.
The reader will recognize this notion as the Weil local system version of equivariant sheaves for the action $n$ of $G\der$ on $G$, as can be found, for example, in \cite{bernstein-lunts:equivariant}*{0.2}.

@@

\begin{lemma}\label{lemma:descent}
The functor $L : \Loc(G\ab)\to \Loc\der(G)$ is an equivalence.
\end{lemma}
\begin{proof}
The quotient $q : G \to G\ab$ is an $G\der$-torsor in the fppf topology by \cite{demazure:SGA3-VIA}*{Thm. 3.2}, and thus a $G\der$-torsor in the fpqc topology.
The lemma is now a result from descent theory, arguing as in \cite{Vistoli:notes}*{Theorem 4.46} for example. 
\end{proof}

Observe that if $(\cs{L},\nu)\in \Loc\der(G)$ then ... \todo{New lemma?}

@@

\subsubsection{Descent}

We may now relate $\CSab(G)$ to $\CS(G\ab)$.
To do so, we use decent along $q : G \to G\ab$.

@@

\begin{proof}
Let $i\der : G\der \to G$ be the kernel of $q:G \to G\ab$ and define $j : G\der \to G\der\times G$ by $j(h) = (h,1)$. 
If $(\cs{L},\mu,\nu)\in \CS\der(G)$ then $j^*(\nu) : \cs{L}\vert_{G\der} \to (\EE)_{G\der}$ is an isomorphism. 
\todo{Address referee's comment here}
This defines a functor by
\[
\begin{array}{rcl}
\CS\der(G) &\to& \CSab(G) \\
(\cs{L},\mu,\nu) &\mapsto& (\cs{L},\mu,j^*(\nu))
\end{array}
\]
on objects and trivially on morphisms.
It is easy to verify that morphisms that commute with $\mu$ and $\nu$ also commute with $\mu$ and $j^*(\nu)$.
This functor is an equivalence; its adjoint is given as follows.
Define $k: G\der\times G\to G\times G$ by $k(h,g) = (i\der(h),g)$.
Then for $(\cs{L},\mu,\beta) \in \CSab(G)$, define $\nu : n^*\cs{L} \to p^*\cs{L}$ by the following diagram.
\[
\begin{tikzcd}
& n^* \cs{L} \arrow[equal]{dl}{m\circ k = n} \arrow[dashed]{r}{\nu} & p^*\cs{L} &  \\ 
k^* m^* \cs{L} \arrow{dr}{k^*(\mu)} &&& (\EE)_{G\der} \boxtimes \cs{L} \arrow[equal]{ul} \\
& k^*(\cs{L}\boxtimes\cs{L}) \arrow[equal]{r} & \cs{L}\vert_{G\der}\boxtimes \cs{L} \arrow{ur}{\beta\boxtimes \id_\cs{L}}
\end{tikzcd}
\]
This defines the functor  $\CS\ab(G) \to \CS\der(G)$, after confirming that morphisms that commute with $\mu$ and $\beta$ also commute with $\mu$ and $\nu$.
\end{proof}

@@

\begin{proof}
Notation as in the proof of Proposition~\ref{prop:CSGabab}.
Consider the following diagram.
\[
\begin{tikzcd}
\CS(G\der) & \arrow{l}[swap]{i\der^*} \CS(G) &\arrow{l}[swap]{q^*} 
\CS(G\ab) \arrow{dl}{q\ab^*} \\
& \CSab(G) \arrow{u}{\text{forget}}  & 
\end{tikzcd}
\]

Now, suppose $(\cs{L},\mu)\in \CS(G)$ and there is an isomorphism $\beta : \cs{L}\vert_{G\der} \to (\EE)_{G\der}$ in $\CS(G\der)$, so that
$(\cs{L},\mu,\beta) \in \CSab(G)$.
By Proposition \ref{prop:CSGabab}, there is some $(\cs{L}\ab,\mu\ab)\in \CS(G\ab)$ with $(\cs{L},\mu,\beta) \iso q\ab^*(\cs{L}\ab,\mu\ab)$.
Applying the forgetful functor $\CS_{G\der}(G)\to \CS(G)$ to this isomorphism, it follows that $(\cs{L},\mu) \iso q^*(\cs{L}\ab,\mu\ab)$ in $\CS(G)$, as desired.

Conversely, suppose $(\cs{L},\mu)\in \CS(G)$ and $(\cs{L},\mu) \iso q^*(\cs{L}\ab,\mu\ab)$ in $\CS(G)$.
Then 
\[
i\der^*(\cs{L},\mu) \iso i\der^*q^*(\cs{L}\ab,\mu\ab)
\]
 in $\CS(G\der)$.
Since $q\circ i\der = 1$, it follows that $\cs{L}\vert_{G\der} \iso (\EE)_{G\der}$ in $\CS(G\der)$.
\end{proof}

@@

We are now in a position to prove that commutative character sheaves on $G$ match perfectly with commutative character sheaves on $G\ab$.
We start with a method that will allow us to situate the diagram in Definition \ref{def:CCS} within $\CSab(G^2)$.

@@

Suppose $G$ is commutative, so $G\der = 1$. 
Suppose $(\cs{L},\mu,\beta)$ is an object in $\CSab(G)$.
Then $\beta : \cs{L}_1\to \EE$ is an isomorphism in $\CS(1)$, which is unique by \cite{cunningham-roe:13a}*{Theorem 3.9}.
Tracing through the construction of $\gamma : m^*\cs{L} \to \theta^*m^*\cs{L}$ from $\beta : \cs{L}_1\to \EE$, we find
that $\gamma : m^*\cs{L} \to \theta^*m^*\cs{L}$ is the canonical isomorphism coming from the equation $m = m \circ \theta$. 
Thus, when $G$ is commutative, Definition~\ref{def:CCS} agrees with Definition~\ref{def:CCScom}.
The next result generalizes this observation.

>>>>

%%%%%%%%%%%% MISCELLANEOUS OPERATORS %%%%%%%%%%%%
\newcommand{\sheafHom}{{\mathscr{H}\hskip-4pt{\it o}\hskip-2pt{\it m}}}
\newcommand{\abs}[1]{{\vert #1 \vert}}
\newcommand{\ceq}{{\, :=\, }}
\newcommand{\tq}{{\ \vert\ }}
\newcommand{\iso}{{\ \cong\ }}
\newcommand{\trFrob}[1]{t_{#1}}
\DeclareMathOperator{\Tr}{Tr}
\newcommand{\TrFrob}[1]{\Tr_{#1}}
%% Limits
\newcommand{\invlim}[1]{\lim\limits_{\overleftarrow{#1}}}
\newcommand{\dirlim}[1]{\lim\limits_{\overrightarrow{#1}}}
\newcommand{\limit}[1]{\mathop{\textsc{lim}}\limits_{#1}}
\newcommand{\colimit}[1]{\mathop{\textsc{colim}}\limits_{#1}}
%% Fonts for quasicharacter sheaves
\newcommand{\cs}[1]{{\mathcal{#1}}}
\newcommand{\gcs}[1]{{\mathcal{\bar #1}}}
\newcommand{\dualgcs}[1]{\gcs{#1}^\dagger}
\newcommand{\dualcs}[1]{\cs{#1}^\dagger}
%% Categories
\newcommand{\CS}{{\mathcal{C\hskip-0.8pt S}}}
\newcommand{\CCS}{{\mathcal{C\hskip-.8pt C\hskip-0.8pt S}}}
\newcommand{\bCS}{{\CS_0}}
%\newcommand{\CSab}{\CS_{\operatorname{ab}}}
\newcommand{\catname}[1]{\normalfont{\textsf{#1}}}
\newcommand{\Sch}[1]{{\catname{Sch}_{/#1}}}
\newcommand{\QCS}{{\mathcal{QC\hskip-0.8pt S}}}
\newcommand{\CSiso}[1]{\CS(#1)_{/\text{iso}}}
\newcommand{\bCSiso}[1]{\bCS(#1)_{/\text{iso}}}
\newcommand{\QCSiso}[1]{\QCS(#1)_{/\text{iso}}}
\newcommand{\CCSiso}[1]{\CCS(#1)_{/\text{iso}}}
%% Labeled items
\makeatletter
\newcommand{\labitem}[2]{
\def\@itemlabel{\textbf{#1}}
\item
\def\@currentlabel{#1}\label{#2}}
\makeatother
%% Shorthand for bars
\renewcommand{\bf}{\bar{f}}
\newcommand{\bg}{{\bar{g}}}
\newcommand{\bm}{\bar{m}}
\newcommand{\bG}{\bar{G}}
\newcommand{\bH}{\bar{H}}
\newcommand{\brho}{{\bar\rho}}
\newcommand{\bx}{{\bar{x}}}
%% Spacing control
\newcommand{\tight}[3]{\hspace{-#1pt}{#2}\hspace{-#3pt}}
\newcommand{\GxG}{\text{$G \tight{1}{\times}{1} G$}}
\newcommand{\bGxG}{\text{$\bar{G} \tight{1}{\times}{1} \bar{G}$}}
\newcommand{\bfxf}{\text{$\bar{f} \tight{1}{\times}{1} \bar{f}$}}
\newcommand{\GxxG}{\text{$G \tight{1}{\times}{1} G$}}
\newcommand{\LxL}{\text{$\gcs{L} \tight{0}{\boxtimes}{0} \gcs{L}$}}

@@

\begin{definition}\label{def:CSab}
Let $\CS\ab(G)$ denote the category of triples $(\cs{L},\mu,\beta)$ where $(\cs{L},\mu) \in \CS(G)$ and
$\beta : \cs{L}\vert_{G\der} \to (\EE)_{G\der}$ is an isomorphism of Weil local systems on $G\der$.
A morphism $(\cs{L},\mu,\beta)\to (\cs{L}',\mu',\beta')$ is a morphism $\alpha : (\cs{L},\mu)\to (\cs{L}',\mu')$
in $\CS(G)$ such that $\beta = \beta' \circ \alpha\vert_{G\der}$.  
\end{definition}

The reason for tracking $\beta$ is that it determines an isomorphism $\gamma : m^*\cs{L} \to \theta^*m^*\cs{L}$ which will replace the $\xi$ of Definition \ref{def:CCScom}, as follows.
Let $i : G \to G$ be inversion and $c : G\times G\to G\der$ be the commutator map, defined by $c(x,y)= xyx^{-1}y^{-1}$.
Both are morphisms of $\Fq$-schemes.
Set $m' = i \circ m \circ \theta$ and let $j\der : G\der\to G$ be inclusion; then $j\der\circ c = m \circ (m \times m')$.
Then $\beta : \cs{L}\vert_{G\der} \to (\EE)_{G\der}$ determines the isomorphism $\gamma' : m^*\cs{L} \otimes \theta^* m^* i^*\cs{L} \to (\EE)_{G\times G}$ by the diagram of isomorphisms below.
\begin{equation}
\begin{tikzcd}
\arrow[equal]{d} c^* (\cs{L}\vert_{G\der}) \arrow{r}{c^*(\beta)} 
	&  c^*((\EE)_{G\der}) \arrow[equal]{d} \\
\arrow{d}[swap]{j\der\circ c = m \circ (m \times m')} c^* j\der^* \cs{L} 
	&   (\EE)_{G\times G} \\
(m \times m')^* m^* \cs{L} \arrow{d}[swap]{(m \times m')^*(\mu)} 
	&  m^*\cs{L} \otimes \theta^* m^* i^* \cs{L}  \arrow[dashed]{u}[swap]{\gamma'} \\
(m \times m')^* (\cs{L} \boxtimes \cs{L}) \arrow[equal]{r} 
	& m^*\cs{L} \otimes (m')^*\cs{L} \arrow{u}[swap]{m' = i\circ m\circ \theta} 
\end{tikzcd}
\end{equation}
In the diagram above, the arrows labeled with equations come from canonical isomorphisms of functors on Weil sheaves derived from the equations; so, for example, the middle left isomorphism comes from $(m\times m')^* m^* \iso c^* j\der^*$ since $j\der\circ c = m \circ (m \times m')$.
Using the monoidal structure of the category of Weil local systems on $G\times G$, the isomorphism $\gamma' : m^*\cs{L} \otimes \theta^* m^* i^*\cs{L} \to (\EE)_{G\times G}$ defines an isomorphism
\[
m^*\cs{L} \to (\theta^* m^* i^*\cs{L})^\vee.
\]
Applying the canonical isomorphisms $(\theta^* m^* i^*\cs{L})^\vee \iso \theta^* m^* i^* (\cs{L}^\vee)$ and $i^*(\cs{L}^\vee) \cong \cs{L}$, this map provides the promised isomorphism
\[
\begin{tikzcd}
\gamma : m^*\cs{L} \arrow{r} & \theta^* m^* \cs{L}.
\end{tikzcd}
\]

\begin{definition}\label{def:CCS}
The category $\CCS(G)$ of commutative character sheaves on $G$ is the full subcategory of $\CS\ab(G)$ consisting of triples $(\cs{L},\mu,\beta)$ such that the following diagram of Weil sheaves on $G \times G$ commutes:
  \[
  \begin{tikzcd}[row sep=30]
   m^*\cs{L} \arrow{d}[swap]{\gamma} \arrow{r}{\mu} & \cs{L}\boxtimes\cs{L} \arrow{d}{\vartheta}\\
   \theta^*(m^*\cs{L}) \arrow{r}{{\theta}^*\mu} &  \theta^*(\cs{L}\boxtimes\cs{L}).
  \end{tikzcd}
  \]
Here $\gamma : m^*\cs{L} \to \theta^* m^* \cs{L}$ is the isomorphism built from $\beta : \cs{L}\vert_{G\der} \to (\EE)_{G\der}$ as above. 
\end{definition}

@@

In this section we give an equivalence of categories between $\CS(G\ab)$ and $\CS\ab(G)$ and use it to describe
the pullback functor $q^* : \CS(G\ab) \to \CS(G)$ in terms of the forgetful functor $\CS\ab(G) \to \CS(G)$, where $q : G\to G\ab$ is the abelianization quotient with kernel $G\der$.
But first, in order to study commutative character sheaves, we need some auxiliary categories.

@@

Let $\Loc(G)$ be the category of Weil local systems on $G$.
Let $\Loc\der(G)$ be the category of $G\der$-equivariant Weil local systems on $G$, whose definition we now recall.
Let $n : G\der\times G\to G$ be the restriction of $m : G\times G\to G$ to $G\der\times G$,
let $p : G\der\times G \to G$ be projection to the second component, and let $s: G \to G\der\times G$ be given by $s(g) = (1,g)$.
Then the quotient $q : G \to G\ab$ is a regular epimorphism of smooth group schemes with kernel pair $(n,p)$.
\[
\begin{tikzcd}
G\der\times G
 \arrow[shift left=1]{r}{n}
  \arrow[shift right=1,swap]{r}{p}
&
G 
\arrow{r}{q}
& 
G\ab
\end{tikzcd}
\]
Consider the morphisms
\[
\begin{tikzcd}
G\der \times G\der \times G 
\arrow[shift left=2]{r}{b_1, b_2, b_3} 
\arrow{r}{}
\arrow[shift right=2]{r}{} 
& G\der \times G 
\arrow[shift left=1]{r}{n}
\arrow[shift right=1]{r}[swap]{p}
 & G 
\end{tikzcd}
\]
defined by 
\begin{align*}
b_1(h_1,h_2,g) &= (h_1h_2,g) \\
b_2(h_1,h_2,g) &= (h_1,h_2g) \\
b_3(h_1,h_2,g) &= (h_2,g).
\end{align*}
Note that
\begin{align*}
n\circ b_1 &= n\circ b_2 \\
n\circ b_3 &= p\circ b_2 \numberthis \label{eqn:bap}\\
p\circ b_1 &= p\circ b_3.
\end{align*}
A $G\der$-equivariant Weil local system on $G$ is a Weil local system $\cs{L}$ on $G$ together with an isomorphism  
\[
\nu : n^*\cs{L} \to p^*\cs{L}
\] 
of Weil local systems on $G\der\times G$ such that 
\begin{equation}\label{E1}
s^*(\nu) = \id_{\cs{L}}
\end{equation}
and such that the following diagram of isomorphisms of Weil local systems on $G\der\times G\der\times G$ commutes. 
\begin{equation}\label{E2}
\begin{tikzcd}
\ &  \arrow{dl}{n\circ b_1 = n\circ b_2}  b_2^*  n^*\, \mathcal{L} \arrow{rr}{b_2^*(\nu)} && b_2^* p^*\, \mathcal{L} \arrow{dr}[swap]{p\circ b_2 = n\circ b_3} & \\
b_1^* n^*\, \mathcal{L} \arrow{dr}{b_1^*(\nu)} &&&&   \arrow{dl}[swap]{b_3^*(\nu)} b_3^* n^*\, \mathcal{L} \\
& b_1^* p^*\, \mathcal{L}  && \arrow{ll}[swap]{p\circ b_3 = p\circ b_1} b_3^* p^*\, \mathcal{L} & 
\end{tikzcd}
\end{equation}
Morphisms of $H$-equivariant Weil local systems $(\cs{L},\nu)\to (\cs{L}',\nu')$ are morphisms of Weil local systems $\alpha: \cs{L}\to \cs{L}'$ for which the diagram
\begin{equation}\label{E3}
\begin{tikzcd}
\arrow{d}[swap]{\nu} n^*\cs{L} \arrow{r}{n^*(\alpha)} & n^*\cs{L}' \arrow{d}{\nu'} \\
p^*\cs{L} \arrow{r}{p^*(\alpha)} & p^*\cs{L}'
\end{tikzcd}
\end{equation}
commutes.
This defines $\Loc\der(G)$, the category of $G\der$-equivariant Weil local systems on $G$.
The reader will recognize this notion as the Weil local system version of equivariant sheaves for the action $n$ of $G\der$ on $G$, as can be found, for example, in \cite{bernstein-lunts:equivariant}*{0.2}.

@@

\begin{lemma}\label{lemma:descent}
Suppose $G$ is a smooth group scheme.  
The functor $L : \Loc(G\ab)\to \Loc\der(G)$ is an equivalence.
\end{lemma}
\begin{proof}
The quotient $q : G \to G\ab$ is an $G\der$-torsor in the fppf topology by \cite{demazure:SGA3-VIA}*{Thm. 3.2}, and thus a $G\der$-torsor in the fpqc topology.
The lemma is now a result from descent theory, arguing as in \cite{Vistoli:notes}*{Theorem 4.46} for example. 
\end{proof}

@@

\subsubsection{Rigidification} \todo{Not sure that's the right subsection title. Other  suggestions?}

We may now relate $\CS\ab(G)$ to $\CS(G\ab)$.
%To do so, we use decent along $q : G \to G\ab$.

@@

\begin{proof}
We begin by defining a functor $\CS\ab(G) \to \CS\der(G)$.
Let $j\der : G\der \to G$ be the kernel of $q:G \to G\ab$ 
Define $k: G\der\times G\to G\times G$ by $k(h,g) = (j\der(h),g)$.
Then for $(\cs{L},\mu,\nu) \in \CS\der(G)$, define $\beta : \cs{L}\vert_{G\der} \to (\EE)_{G\der}$ by the following diagram.
\[
\begin{tikzcd}
& n^* \cs{L} \arrow[equal]{dl}{m\circ k = n} \arrow{r}{\nu} & p^*\cs{L} &  \\ 
k^* m^* \cs{L} \arrow{dr}{k^*(\mu)} &&& (\EE)_{G\der} \boxtimes \cs{L} \arrow[equal]{ul} \\
& k^*(\cs{L}\boxtimes\cs{L}) \arrow[equal]{r} & \cs{L}\vert_{G\der}\boxtimes \cs{L} \arrow{ur}{\beta\boxtimes \id_\cs{L}}
\end{tikzcd}
\]
Then $\beta = r\der \circ j^*(\nu)$, where $r\der : (\cs{L}_1)_{G\der} \to (\EE)_{G\der}$ is the isomorphism of constant local systems  determined by the rigidification $r: \cs{L}_1 \to \EE$ of $\cs{L}$ determined by $\mu$  (see \cite{cunningham-roe:13a}*{Remark~1.11}).
To show that $(\cs{L},\mu,\beta) \in \CS\ab(G)$, we must see that $\beta$ is a morphism in $\CS(G\der)$.
By Proposition~\ref{prop:CSGabab}, $\cs{L} = q^*\cs{L}\ab$.

This defines the functor  $\CS\ab(G) \to \CS\der(G)$, after confirming that morphisms that commute with $\mu$ and $\beta$ also commute with $\mu$ and $\nu$.

and define $j : G\der \to G\der\times G$ by $j(h) = (h,1)$.

If $(\cs{L},\mu,\nu)\in \CS\der(G)$ then $j^*(\nu) : \cs{L}\vert_{G\der} \to (\EE)_{G\der}$ is an isomorphism. 
\todo{Address referee's comment here}
This defines a functor by
\[
\begin{array}{rcl}
\CS\der(G) &\to& \CS\ab(G) \\
(\cs{L},\mu,\nu) &\mapsto& (\cs{L},\mu,j^*(\nu))
\end{array}
\]
on objects and trivially on morphisms.
It is easy to verify that morphisms that commute with $\mu$ and $\nu$ also commute with $\mu$ and $j^*(\nu)$.
This functor is an equivalence; its adjoint is given as follows.

@@

\begin{proof}
Notation as in the proof of Proposition~\ref{prop:CSGabab}.
Consider the following diagram.
\[
\begin{tikzcd}
\CS(G\der) & \arrow{l}[swap]{j\der^*} \CS(G) &\arrow{l}[swap]{q^*} 
\CS(G\ab) \arrow{dl}{q\ab^*} \\
& \CS\ab(G) \arrow{u}{\text{forget}}  & 
\end{tikzcd}
\]

Now, suppose $(\cs{L},\mu)\in \CS(G)$ and there is an isomorphism $\beta : \cs{L}\vert_{G\der} \to (\EE)_{G\der}$ so that
$(\cs{L},\mu,\beta) \in \CS\ab(G)$.
By Proposition \ref{prop:CSGabab}, there is some $(\cs{L}\ab,\mu\ab)\in \CS(G\ab)$ with $(\cs{L},\mu,\beta) \iso q\ab^*(\cs{L}\ab,\mu\ab)$.
Applying the forgetful functor $\CS_{G\der}(G)\to \CS(G)$ to this isomorphism, it follows that $(\cs{L},\mu) \iso q^*(\cs{L}\ab,\mu\ab)$ in $\CS(G)$, as desired.

Conversely, suppose $(\cs{L},\mu)\in \CS(G)$ and $(\cs{L},\mu) \iso q^*(\cs{L}\ab,\mu\ab)$ in $\CS(G)$.
Then 
\[
j\der^*(\cs{L},\mu) \iso j\der^*q^*(\cs{L}\ab,\mu\ab)
\]
 in $\CS(G\der)$.
Since $q\circ j\der = 1$, it follows that $\cs{L}\vert_{G\der} \iso (\EE)_{G\der}$.
\end{proof}

@@

We are now in a position to prove that commutative character sheaves on $G$ match perfectly with commutative character sheaves on $G\ab$.
We start with a method that will allow us to situate the diagram in Definition \ref{def:CCS} within $\CS\ab(G^2)$.

@@

Suppose $G$ is commutative, so $G\der = 1$. 
Suppose $(\cs{L},\mu,\beta)$ is an object in $\CS\ab(G)$.
Then $\beta : \cs{L}_1\to \EE$ is an isomorphism in $\CS(1)$, which is unique by \cite{cunningham-roe:13a}*{Theorem 3.9}.
Tracing through the construction of $\gamma : m^*\cs{L} \to \theta^*m^*\cs{L}$ from $\beta : \cs{L}_1\to \EE$, we find
that $\gamma : m^*\cs{L} \to \theta^*m^*\cs{L}$ is the canonical isomorphism coming from the equation $m = m \circ \theta$. 
Thus, when $G$ is commutative, Definition~\ref{def:CCS} agrees with Definition~\ref{def:CCScom}.
The next result generalizes this observation.
